Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-02-15 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: explained mapping in concrete example; terse

\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath,longbibliography]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
% \usepackage{txfonts}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
\usepackage{mathrsfs}
\usepackage{makecell}
\usepackage{microtype}

@@

Let us now establish the connection between theory-agnostic framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.
%
We begin by expanding $\omega$ and $\tau$, given by Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\begin{align}
M \omega_{220} &= \gamma (\delta\omega^{(0)} \omega^{(0)} + \chi \delta\omega^{(1)} \omega^{(1)})
+ \sum_{j=0}^{N} \chi^{j} \omega^{j}\,,
\\
\tau_{220}/M   &= \gamma (\delta\tau^{(0)} \tau^{(0)} + \chi \delta\tau^{(1)} \tau^{(1)})
+ \sum_{j=0}^{N} \chi^{j} \tau^{j}\,,
\end{align}
\label{eq:qnm_nongr_iso}
\end{subequations}
%
where we are already focusing on the dominant $\ell = m = 2$, $n=0$ QNM and
pulled out of the sum all beyond-GR corrections, rescriting ourselves to the nonspinning ($j=0$) and linear-order in spin ($j=1$)
corrections to the Kerr QNMs, encoded in the coefficients $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$.

Just as in GR where the coefficients $\omega^{j}$, $\tau^{j}$ are determined by
comparison against the Kerr QNMs, a theory-specific calculation predicts the
numerical values of $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$
%
This means that using the $\pSEOB$ waveform model with theory-specific informed
QNMs we reduce the parameters space of beyond-GR parameters to one, i.e., to
only $\gamma$, which in turn contains information about the length scale $\ell$.

@@

>>>>

\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath,longbibliography]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{txfonts}
\usepackage{natbib}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
\usepackage{mathrsfs}
\usepackage{makecell}
\usepackage{microtype}

@@

Let us now establish the connection between theory-agnostic framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.
%
We begin by expanding $\omega$ and $\tau$, given by Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\begin{align}
M \omega_{220} &= \gamma \left( \delta\omega^{(0)} \omega^{(0)} + \chi \delta\omega^{(1)} \omega^{(1)} \right)
+ \sum_{j=0}^{N} \chi^{j} \omega^{j}\,,
\\
\tau_{220}/M   &= \gamma \left( \delta\tau^{(0)} \tau^{(0)} + \chi \delta\tau^{(1)} \tau^{(1)} \right)
+ \sum_{j=0}^{N} \chi^{j} \tau^{j}\,,
\end{align}
\label{eq:qnm_nongr_iso}
\end{subequations}
%
where we are already focusing on the dominant $\ell = m = 2$, $n=0$ QNM and
pulled out of the sum all beyond-GR corrections, rescriting ourselves to the nonspinning ($j=0$) and linear-order in spin ($j=1$)
corrections to the Kerr QNMs, encoded in the coefficients $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$.

Just as in GR where the coefficients $\omega^{j}$, $\tau^{j}$ are determined by
comparison against the Kerr QNMs, a theory-specific calculation predicts the
numerical values of $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$
%
This means that using the $\pSEOB$ waveform model with theory-specific informed
QNMs we reduce the parameters space of beyond-GR parameters to one, i.e., to
only $\gamma$, which in turn contains information about the length scale $\ell$.
%
It is important to emphasize that our procedure is different from that of~\cite{Carullo:2021dui}
which, for a given value of $p$, varied all $\ell$, $\delta\omega^{(i)}$ and $\delta\tau^{(i)}$ parameters.
%
% As we will see later, our theory-informed approach can result to stronger constraints to certain modified ....

@@

% ---------------------------
% CONCRETE EXAMPLE OF MAPPING
% ---------------------------
Let us show one concrete example of how this mapping is done.
%
In Ref.~\cite{Wagle:2021tam} the QNMs of slow-rotating BHs in dCS gravity were calculated
and it was found that the axial gravitational modes have their damping time increased as we
increase the Chern-Simons coupling $\ell_{\rm CS}$ at fixed spin $\chi$ of the BH.
%
Having selected this mode we can now obtain $\delta\omega^{(i)}$ and $\delta\tau^{(i)}$ as follows.
%
Using the fitting formula Eq.~(54a) of~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M \omega^{\rm CS}_{220} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi)^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part ${\rm Im}(\sigma_{220}) = \omega^{\rm CS}_{220} = 1 / \tau^{\rm CS}_{220}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm CS}^{4} / (M^4 \kappa)$ [such that $\kappa \zeta = (\ell_{\rm CS} / M)^4$]
and $c_{i}$ are fitting coefficients.
%
In the context of a binary BH merger, $M$ is interpreted as the remnant mass $\Mf$.

We can now expand Eq.~\eqref{eq:omega_dsc_eg} to leading-orders in $\chi$ and $\ell_{\rm CS}$, and
gather the terms proportional to $\ell_{\rm CS}$.
%
Doing so, we find,
%
\begin{align}
    \Mf \, \omega^{\rm CS}_{220} &=
    \left[ 0.3722 + 1.1945 (\ell_{\rm CS} / \Mf)^4 \right]
    \nonumber \\
    &\quad + \left[0.1861 + 5.1828 (\ell_{\rm CS} / \Mf)^4 \right] \, \chi\,,
\end{align}
%
where we made use of the numerical value of the coefficients $c_{i}$ and,
reassuringly, the nonrotating and linear-in-spin coefficients above agree with those
of~\cite{Maselli:2019mjd} to xx\%.
%
We now isolate the dCS corrections and compare against Eq.~\eqref{eq:qnm_nongr_iso}.
%
Doing so, we can make the identifications,
%
\begin{equation}
p_{\rm cs} = 4\,, \quad \delta \omega^{(0)}_{\rm cs} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm cs} = 41.199\,.
\label{eq:cs_omega_coefs}
\end{equation}
%
The same procedure can be made for $\tau^{(i)}$, which results in,
%
\begin{equation}
\delta \tau^{(0)}_{\rm cs} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm cs} = 794.66\,.
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed parameters beyond-GR parameters in the ringdown of the \pSEOB waveform model.
%
As we said before, the only free beyond-GR parameter left is now $\ell_{\rm CS}$.
